### 2021-01-10 tex/paper.tex
Add more findings and conclusion.

\begin{itemize}
  \item ease of use
  \item customisability
  \item scalability
  \item suitability for mission-critical applications
\end{itemize}

@@

The Corteza-Low-Code platform is spreadsheet based, making is very easy to use. Creating a simple TODO List application was possible in less than 10 minutes and should be possible in less than 5 minutes once you are familiar with the platform. Apart from choosing the order of predefined building blocks, customising the layout of an application is hindered by the fact that it is spreadsheet based. The selection of pre-built blocks is very limited, as is the selection of color schemes. In terms of scalabilty, creating complex applications will be difficult using the small selection of building blocks. Process management applications which don't rely on many different UI elements, like the included Corteza CRM, can definitely be implemented in a scalable manner. Given the overall simplicity of the Corteza platform, it is certainly suitable for creating mission-critical applications.

@@

Creating an application Oracle APEX requires basic SQL knowledge and the page designer is quite complex. Nevertheless, we were able to create our TODO List application in less than two hours by reviewing a sample application. It is also possible to completely generate the basic structure of an application using a wizard by importing data from a file, making the initial setup even easier. For further development, there is a small learning curve for getting familiar with the interface. The default theme offers three different menu types and four colour schemes and there is an option to use a custom theme. The page editor offers all basic building blocks in addition to some specialised ones, the ability to execute custom actions on page/button events and advanced options for each component. Given the flexibility and advanced options of the page designer, it is definitely possible to create scalable applications which can also be used for mission-critical usage.

@@

The Simplifier application editor offers an extensive list of UI components for composing an application with advanced options for each component, which is of course associated with a small learning curve to get familiar with the interface. A different theme can be uploaded manually. All things considered, this make an application very customisable and the platform itself moderately easy to use. Adding custom functionality can be done easily by adding so called “stories” using a graphical editor as well, making it very scalable since these “stories” can then be composed by creating custom events. Leaving aside the technical difficulties we had during the creation of our TODO List application, we still think it is suitable for creating mission-critical applications, given the commercial support would most likely be able to sort out these issues.

@@

% TODO

\section{Findings & Possible Future Work}
\label{sec:findings}

@@

% TODO: Difference between open-source/commercial, old/new, etc.

@@

>>>>

\begin{itemize}
  \item ease of use
  \item customisability
  \item portability
  \item scalability
  \item suitability for mission-critical applications
\end{itemize}

@@

The Corteza-Low-Code platform is spreadsheet based, making is very easy to use. Creating a simple TODO List application was possible in less than 10 minutes and should be possible in less than 5 minutes once you are familiar with the platform. Apart from choosing the order of predefined building blocks, customising the layout of an application is hindered by the fact that it is spreadsheet based. The selection of pre-built blocks is very limited, as is the selection of color schemes. Corteza uses a MySQL database in the backend, making at least the stored data very portable. The created applications are virtually unportable since they are tightly integrated into the Corteza platform, custom scripts are at least partially portable since they are written in JavaScript. In terms of scalabilty, creating complex applications will be difficult using the small selection of building blocks. Process management applications which don't rely on many different UI elements, like the included Corteza CRM, can definitely be implemented in a scalable manner. Given the overall simplicity of the Corteza platform, it is certainly suitable for creating mission-critical applications.

@@

Creating an application Oracle APEX requires basic SQL knowledge and the page designer is quite complex. Nevertheless, we were able to create our TODO List application in less than two hours by reviewing a sample application. It is also possible to completely generate the basic structure of an application using a wizard by importing data from a file, making the initial setup even easier. For further development, there is a small learning curve for getting familiar with the interface. The default theme offers three different menu types and four colour schemes and there is an option to use a custom theme. The page editor offers all basic building blocks in addition to some specialised ones, the ability to execute custom actions on page/button events and advanced options for each component. Given the flexibility and advanced options of the page designer, it is definitely possible to create scalable applications which can also be used for mission-critical usage. Oracle APEX uses their propietary Oracle Database for the backend and custom logic has to be written in PL/SQL, a propietary programming language also by Oracle. This makes an Oracle APEX application completely unportable and vendor lock-in is a definate issue.

@@

The Simplifier application editor offers an extensive list of UI components for composing an application with advanced options for each component, which is of course associated with a small learning curve to get familiar with the interface. A different theme can be uploaded manually. All things considered, this make an application very customisable and the platform itself moderately easy to use. Adding custom functionality can be done easily by adding so called “stories” using a graphical editor as well, making it very scalable since these “stories” can then be composed by creating custom events. Leaving aside the technical difficulties we had during the creation of our TODO List application, we still think it is suitable for creating mission-critical applications, given the commercial support would most likely be able to sort out these issues. A Simplifier application relies heavily on “stories”, which are stored in a propietary format and there for not portable. The data, however, is highly portable since it is stored on external databases via connectors and many different databases are supported.

@@

Mendix Studio offers many building-blocks and widgets to choose from and also offers a marketplace where developers can publish custom ones, making the platform very flexible. The color scheme of an application can also be completely customised and the marketplace also offers the ability to use custom themes. Due to the fact that the data and building blocks are tightly coupled, it is very easy to create basic application logic. Application data can be downloaded from backups in SQLite format and external data sources can be installed form the marketplace, making the data very portable. The application itself it tightly coupled to the Mendix platform and not at all portable. Given the vast amount of building-blocks available (counting those available from the marketplace), Mendix is certainly suitable for building scalable and mission-critical applications.

\section{Findings \& Possible Future Work}
\label{sec:findings}

@@

Another thing that stood out is the user-interface design of the low-code platforms. There is a clear trend towards a more streamlined user interface in the more mature products like Oracle APEX and Mendix. The user interface of, say Corteza, while simple, is not as thought-through and the Simplifier interface, while functional, looks like a generic UI prototype. Whether this makes a platform more usable than another lies of course in the eye of the beholder.

A common point of critique of low-code platforms is the lack of portability. According to our evaluation, this point is definitely valid. Of the platforms we evaluated, none of the applications are portable; at most, the stored data is easily portable by using a common database backend like MySQL. There is no standard definition of what a low-code platform looks like let alone any standard defining low-code building blocks which could be used to make porting an application from one platform to another possible. Given that many low-code platforms are commercial products, it is also doubtful if there is any incentive to make it easy for customers to switch platforms.

@@

Taking the assessments in \cref{sec:assessment_of_tools} into account, low-code platforms are a valid alternative to model-driven development. Some caveats apply however: By their very nature, low-code platforms will not be as flexible, since with model-driven architecture, developers can take advantage of the full feature set of the software stack and customise all aspects of it, while low-code platforms are limited in the ways they can be extended. This is of course highly dependent on the platform, which means one cannot argue for or against low-code in all cases. The choice between low-code and model-driven architecture cannot be made without evaluating multiple different low-code platforms and assessing whether they can satisfy the requirements for a particular project, similar to choosing a programming language or framework that is most suitable for a particular task.
